---
title: "Assign03"
author: "Shivam"
output: pdf_document
date: "2023-02-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#if(!is.null(dev.list())) dev.off()
#cat("\014")
#rm(list=ls())
options(scipen=9)
```

```{r}
if(!require(pastecs)){install.packages("pastecs")}
library("pastecs")

if(!require(lattice)){install.packages("lattice")}
library("lattice")
```

```{r}
setwd("C:/Users/holys/OneDrive/Desktop/Data Analytics,Mathamatics,Algor/Assign03")

test_SJ <- read.table("PROG8430-23W-Assign03.txt", sep=",", header=TRUE)
    test_SJ <- as.data.frame(test_SJ)
    head(test_SJ)
str(test_SJ)

```
********************************************************************************
1. Data Transformation


    1. Rename all variables with your initials appended (just as was done in 
        assignment 1)
      
```{r}
colnames(test_SJ) <- paste(colnames(test_SJ), "SJ", sep = "_")  # renaming all the 
                                                                # variables and adding
                                                                # my initials SJ
  head(test_SJ)

```
      
     
    2. Standardize all of the variables using either of the two functions 
     demonstrated in class. Describe why you chose the method you did.
     
    -> Before standardization of any of the variable, first lets plot boxplot to 
      observe the distribution of data.
```{r}

 par(mfrow=c(2,2))                                #defining format of the boxplot

for (i in 1:ncol(test_SJ)) {                      # Generating box plot for each 
                                                  # variable in th data set using 
                                                  # for loop
  if (is.numeric(test_SJ[,i])) {
      boxplot(test_SJ[i], main= names(test_SJ)[i],
              horizontal=TRUE, pch=10)
  }
}


par(mfrow=c(1,1))

```

    ->  By observing and analyzing, it is interpreted that most of the data is not abnormal
        and very few vaiables out of the total, has a very little outliers.
        In this case as the data has a very few outliers,I choose to perform min-max 
        standardization formula for different variables. I chose this method as it preserves
        the range of data and this method is more suitable here as we do not want to get the
        data in N(0,1), rather we want it over (0,1).
    
    
```{r}
#standardization function
Std_SJ <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))          
}
```


```{r}

test_SJ$Food_Std_SJ  <- Std_SJ(test_SJ$Food_SJ)   # standardizing each variable by
                                                  # passing it standardization function 
test_SJ$Enter_Std_SJ <- Std_SJ(test_SJ$Enter_SJ)  # created above


test_SJ$Edu_Std_SJ   <- Std_SJ(test_SJ$Edu_SJ)

test_SJ$Trans_Std_SJ <- Std_SJ(test_SJ$Trans_SJ)

test_SJ$Work_Std_SJ  <- Std_SJ(test_SJ$Work_SJ)

test_SJ$House_Std_SJ <- Std_SJ(test_SJ$House_SJ)

test_SJ$Oth_Std_SJ   <- Std_SJ(test_SJ$Oth_SJ)

head(test_SJ)
```

2 Descriptive Data Analysis

    1. Create graphical summaries of the data (as demonstrated in class: 
       boxplots, histograms or density plots) and comment on any 
       observations you make.
      -> To graphically summarize the data, lets generate box plot:

```{r}

par(mfrow=c(2,2))

for (i in 1:ncol(test_SJ)) {
  if (is.numeric(test_SJ[,i])) {
      boxplot(test_SJ[i], main=names(test_SJ)[i],
              horizontal=TRUE, pch=10)
  }
}                              # plotting the box plot of every variable 
                               # including the box plots of standardized 
par(mfrow=c(1,1))              # variables as well

```
   -> From the box plots, it is observed that all the variable have varried ranges 
      and very few of them have a little outliers including Entertainment and
      transportation.
   
   
```{r}
# plotting density plots for graphically
#analyzing data in detailed way of standardized variables
densityplot( ~ test_SJ$Food_Std_SJ, pch=6)   
densityplot( ~ test_SJ$Enter_Std_SJ, pch=6)
densityplot( ~ test_SJ$Edu_Std_SJ, pch=6)
densityplot( ~ test_SJ$Trans_Std_SJ, pch=6)
densityplot( ~ test_SJ$Work_Std_SJ, pch=6)
densityplot( ~ test_SJ$House_Std_SJ, pch=6)
densityplot( ~ test_SJ$Oth_Std_SJ, pch=6)

```
  -> By observing the density plots, one can infer that in each the plots most 
    of the data point are in the range and very few outliers. 
  
  
********************************************************************************
  
  3 Clustering
    Using the K-Means procedure as demonstrated in class, create 
    clusters with k=2,3,4,5,6,7.
    You will be using only two variables as your centroids (House and 
    Food)
    
    -> We have already standardized two variables House and Food as Food_Std_SJ
        and House_Std_SJ.
```{r}
# these are the two variables Food and House in standardized form

centroids_for_plotting_SJ <- c(test_SJ$Food_Std_SJ,test_SJ$House_Std_SJ)

head(centroids_for_plotting_SJ)
```
      
      
    1. Create segmentation/cluster schemes for k=2,3,4,5,6,7. 
    ->  Now I made variables below for generating elbow chart to identify and
        choose the value of K to perform K means clustering.
       

```{r}

# Creating Variable for Elbow Chart
# Trying for 2 to 7 Clusters
maxk_SJ <- 7
nk_SJ <- c(2:maxk_SJ)
wss_SJ <- rep(0,maxk_SJ-1)

```
    
    
    
```{r}
# As asked in the question,firstly I created clusters keeping the value of k=2

#Set Number of Clusters
k=2

Clstr_SJ <- kmeans(test_SJ[,c(8,13)], iter.max=10, centers=k, nstart=10)
Clstr_SJ$size
Clstr_SJ$centers
Clstr_SJ$betweenss/Clstr_SJ$totss

test_SJ$cluster <- factor(Clstr_SJ$cluster)   # Adding Cluster tags to variables
head(test_SJ)

centers_SJ <- data.frame(cluster=factor(1:k), Clstr_SJ$centers)

wss_SJ[k-1] <- Clstr_SJ$tot.withinss

```

```{r}
#Set Number of Clusters by putting the value of k as 3.
k=3

Clstr_SJ <- kmeans(test_SJ[,c(8,13)], iter.max=10, centers=k, nstart=10)
Clstr_SJ$size
Clstr_SJ$centers
Clstr_SJ$betweenss/Clstr_SJ$totss

test_SJ$cluster <- factor(Clstr_SJ$cluster)   # Adding Cluster tags to variables
head(test_SJ)

centers_SJ <- data.frame(cluster=factor(1:k), Clstr_SJ$centers)

wss_SJ[k-1] <- Clstr_SJ$tot.withinss

```
      
```{r}
#Setting the value of k as 4 and creating clusters
k=4

Clstr_SJ <- kmeans(test_SJ[,c(8,13)], iter.max=10, centers=k, nstart=10)
Clstr_SJ$size
Clstr_SJ$centers
Clstr_SJ$betweenss/Clstr_SJ$totss

test_SJ$cluster <- factor(Clstr_SJ$cluster)   # Adding Cluster tags to variables
head(test_SJ)

centers_SJ <- data.frame(cluster=factor(1:k), Clstr_SJ$centers)

wss_SJ[k-1] <- Clstr_SJ$tot.withinss

```
      
```{r}
# For value of k as 5 
k=5

Clstr_SJ <- kmeans(test_SJ[,c(8,13)], iter.max=10, centers=k, nstart=10)
Clstr_SJ$size
Clstr_SJ$centers
Clstr_SJ$betweenss/Clstr_SJ$totss


test_SJ$cluster <- factor(Clstr_SJ$cluster)   # Adding Cluster tags to variables
head(test_SJ)

centers_SJ <- data.frame(cluster=factor(1:k), Clstr_SJ$centers)

wss_SJ[k-1] <- Clstr_SJ$tot.withinss

```

```{r}
# For value of k=6
k=6

Clstr_SJ <- kmeans(test_SJ[,c(8,13)], iter.max=10, centers=k, nstart=10)
Clstr_SJ$size
Clstr_SJ$centers
Clstr_SJ$betweenss/Clstr_SJ$totss

test_SJ$cluster <- factor(Clstr_SJ$cluster)   # Adding Cluster tags to variables
head(test_SJ)

centers_SJ <- data.frame(cluster=factor(1:k), Clstr_SJ$centers)

wss_SJ[k-1] <- Clstr_SJ$tot.withinss

```
      
```{r}
#Set Number of Clusters as k= 7
k=7

Clstr_SJ <- kmeans(test_SJ[,c(8,13)], iter.max=10, centers=k, nstart=10)
Clstr_SJ$size
Clstr_SJ$centers
Clstr_SJ$betweenss/Clstr_SJ$totss

test_SJ$cluster <- factor(Clstr_SJ$cluster)   # Adding Cluster tags to variables
head(test_SJ)

centers_SJ <- data.frame(cluster=factor(1:k), Clstr_SJ$centers)

wss_SJ[k-1] <- Clstr_SJ$tot.withinss

```

      2. Create the WSS plots as demonstrated in class and select a     
         suitable k value based on the “elbow”. [NOTE – Use the code 
         that I provided to do this. Using other functions will yield 
         different results.
         
        -> As thought in the class, and taking the reference form the example provided,
          lets make an elbow plot to choose the value of k.
         
         
```{r}
# Plotting the elbow chart after trying the value of k=2,3,4,5,6,7
plot(2:maxk_SJ, wss_SJ,
     type="b", pch = 19, frame = FALSE,   
     main="Elbow Chart for Clusters",
     xlab="Values of K",
     ylab="Total Within-Clusters SS",
     ylim=c(0,max(wss_SJ)))


```
     -> As observed from the elbow plot,it is clearly observed that the bend is 
      observed when the value of k is 3. Hence, value of k=3.
      
********************************************************************************

  4. Evaluation of Clusters
      1. Based on the “k” chosen above, create a scatter plot showing the 
        clusters and colour-coded data points for each of “k-1”, “k”, “k+1”. For 
        example, if you think the “elbow” is at k=4 create the charts for k=3, 
        k=4 and k=5.
        
        -> As value of k is 3, lets make scatter plot for k=2, 3 and 4.
        
```{r}
k=2

Clstr_SJ <- kmeans(test_SJ[,c(8,13)], iter.max=10, centers=k, nstart=10)
Clstr_SJ$size
Clstr_SJ$centers
Clstr_SJ$betweenss/Clstr_SJ$totss


test_SJ$cluster <- factor(Clstr_SJ$cluster)   
head(test_SJ)

centers_SJ <- data.frame(cluster=factor(1:k), Clstr_SJ$centers)

wss_SJ[k-1] <- Clstr_SJ$tot.withinss

plot(test_SJ$Food_Std_SJ, test_SJ$House_Std_SJ, #ploting scatter plot for analyzing
     col=test_SJ$cluster, 
      main=" When K=3 ",
     xlab="Food ",
     ylab="House",
     pch=as.numeric(test_SJ$cluster)) # when k= 2
points(centers_SJ$Food_Std_SJ, centers_SJ$House_Std_SJ,
       col=centers_SJ$cluster, pch=as.numeric(centers_SJ$cluster),
       cex=3, lwd=3)



```
        
        
```{r}
k=3

Clstr_SJ <- kmeans(test_SJ[,c(8,13)], iter.max=10, centers=k, nstart=10)
Clstr_SJ$size
Clstr_SJ$centers
Clstr_SJ$betweenss/Clstr_SJ$totss


test_SJ$cluster <- factor(Clstr_SJ$cluster)   
head(test_SJ)

centers_SJ <- data.frame(cluster=factor(1:k), Clstr_SJ$centers)

wss_SJ[k-1] <- Clstr_SJ$tot.withinss

plot(test_SJ$Food_Std_SJ, test_SJ$House_Std_SJ, #ploting scatter plot for analyzing
     col=test_SJ$cluster, 
      main=" When K=3 ",
     xlab="Food ",
     ylab="House",
     pch=as.numeric(test_SJ$cluster)) # when k=3
points(centers_SJ$Food_Std_SJ, centers_SJ$House_Std_SJ,
       col=centers_SJ$cluster, pch=as.numeric(centers_SJ$cluster),
       cex=3, lwd=3)



```

```{r}
k=4

Clstr_SJ <- kmeans(test_SJ[,c(8,13)], iter.max=10, centers=k, nstart=10)
Clstr_SJ$size
Clstr_SJ$centers
Clstr_SJ$betweenss/Clstr_SJ$totss

test_SJ$cluster <- factor(Clstr_SJ$cluster)   
head(test_SJ)

centers_SJ <- data.frame(cluster=factor(1:k), Clstr_SJ$centers)

wss_SJ[k-1] <- Clstr_SJ$tot.withinss

plot(test_SJ$Food_Std_SJ, test_SJ$House_Std_SJ, #ploting scatter plot for analyzing
     col=test_SJ$cluster, 
      main=" When K= 4",
     xlab="Food ",
     ylab="House",
     pch=as.numeric(test_SJ$cluster)) # when k= 4
points(centers_SJ$Food_Std_SJ, centers_SJ$House_Std_SJ,
       col=centers_SJ$cluster, pch=as.numeric(centers_SJ$cluster),
       cex=3, lwd=3)
```
        


      2. Based on the WSS plot (3.2) and the charts (4.1) choose one set of 
        clusters that best describes the data.
      -> By observing the elbow chart and then plotting the scatter plot, the cluster 
          which is having lowest values for houses and food, is the one in which
          the points are most thighly bound. So, it describes data the best when 
          the value of k is 3.
          
      
        
      3. Create summary tables for the segmentation/clustering scheme 
        (selected in step 4.2).
        
```{r}

SummClusters_SJ <- aggregate(cbind(Food_SJ,Enter_SJ ,Edu_SJ,Trans_SJ,
                           Work_SJ,House_SJ,Oth_SJ) ~ cluster, test_SJ,
                           FUN=mean)
SummClusters_SJ


```

         
         4. Create suitable descriptive names for each cluster.
         
        -> By analyzing the three cluster s while putting the value of k as 3,
            the relevent titles of each of them as as follows:
                Cluster 1: Food and Housing Insecurity/ Lower Class / Poor
                          - as observed, people having very less food and 
                            very little house or no houses, this is the category
                            or cluster of people belonging to very poor financial
                            background.
                Cluster 2: Basic needs met/ Middle class/ Modetate 
                          - this is the cluster which belongs to the middle clas
                          and upper middle class pople having gtheir basic 
                          neccecities fulfiled.
                Cluster 3: Food and Housing Security/ High class / Rich
                          - this is the cluster representing the priviledged people
                           who are financially free.
              
      
        5. Suggest possible uses for this clustering scheme.

         -> By this clustering, one group unorganized and unlabelled under 
            differnet clusters, depending on the value of k decided. 
           By using the clustering sceme the government can analyze and make 
           dicisions for the betterment of any of the group identifiend. 
           
          According to the clustering, one can know where more number of houses
          are there, this can help in real estate market analysis.The urabn planning
          of the city can also be done on the basis of the inffered data.
          
          Planning of food supply, and schemes for food in low prizes can be done
          by the government by the data. 
          
         
##############################################################################################################         
         
